---
title: 'Bios 6301: Assignment 7'
author: 'Haley Yaremych'
output: pdf_document
---

*Due Thursday, 04 November, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

40 points total.

Submit a single knitr file (named `homework7.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework7.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**21 points**

Use the following code to generate data for patients with repeated measures of A1C (a test for levels of blood glucose).

```{r}
genData <- function(n) {
    if(exists(".Random.seed", envir = .GlobalEnv)) {
        save.seed <- get(".Random.seed", envir= .GlobalEnv)
        on.exit(assign(".Random.seed", save.seed, envir = .GlobalEnv))
    } else {
        on.exit(rm(".Random.seed", envir = .GlobalEnv))
    }
    set.seed(n)
    subj <- ceiling(n / 10)
    id <- sample(subj, n, replace=TRUE)
    times <- as.integer(difftime(as.POSIXct("2005-01-01"), as.POSIXct("2000-01-01"), units='secs'))
    dt <- as.POSIXct(sample(times, n), origin='2000-01-01')
    mu <- runif(subj, 4, 10)
    a1c <- unsplit(mapply(rnorm, tabulate(id), mu, SIMPLIFY=FALSE), id)
    data.frame(id, dt, a1c)
}
x <- genData(500)
```

Perform the following manipulations: (3 points each)

1. Order the data set by `id` and `dt`.
```{r}
ord = x[order(x[,'id'], x[,'dt']),]
```

2. For each `id`, determine if there is more than a one year gap in between observations.  Add a new row at the one year mark, with the `a1c` value set to missing.  A two year gap would require two new rows, and so forth.
```{r}
for (i in 1:length(unique(ord$id))){
  idx = ord[which(ord$id==i),]
  # differences in seconds
  diffs = diff(as.numeric(idx$dt))
  # differences in days
  diffdays = diffs/86400
  
  which = which(diffdays>365)
  
  
  # couldn't figure this one out, sorry!

  
}
```


3. Create a new column `visit`.  For each `id`, add the visit number.  This should be 1 to `n` where `n` is the number of observations for an individual.  This should include the observations created with missing a1c values.
```{r}
ord$visit = numeric(dim(ord)[1])
for (i in 1:length(unique(ord$id))){
  n = nrow(ord[which(ord$id==i),])
  ord[which(ord$id==i),"visit"] = seq(n)
}
```

4. For each `id`, replace missing values with the mean `a1c` value for that individual.
```{r}
for (i in 1:length(unique(ord$id))){
  meana1c = mean(ord[which(ord$id==i),"a1c"], na.rm=TRUE)
  
  ord[which(ord$id==i),is.na("a1c")] = meana1c
}  
```

5. Print mean `a1c` for each `id`.
```{r}
tapply(ord$a1c, ord$id, mean)
```
6. Print total number of visits for each `id`.
```{r}
tapply(ord$visit, ord$id, max)
```
7. Print the observations for `id = 15`.
```{r}
ord[which(ord$id==15),]
```

### Question 2 ###

**16 points**

Install the `lexicon` package. Load the `sw_fry_1000` vector, which contains 1,000 common words.

```{r}
data('sw_fry_1000', package = 'lexicon')
head(sw_fry_1000)
```

1. Remove all non-alphabetical characters and make all characters lowercase. Save the result as `a`.

```{r}
for (i in 1:1000){
  sw_fry_1000[i] = tolower(sw_fry_1000[i])
  sw_fry_1000[i] = gsub("[^a-z]", "", sw_fry_1000[i])
}
a = sw_fry_1000
```

Use vector `a` for the following questions. (2 points each)

2. How many words contain the string "ar"?

```{r}
length(grep("ar", a))
```

3. Find a six-letter word that starts with "l" and ends with "r".

```{r}
grep("^l.{4}r$", a, value=TRUE)
```

4. Return all words that start with "col" or end with "eck".

```{r}
grep("^col|eck$", a, value=TRUE)
```

5. Find the number of words that contain 4 or more adjacent consonants. Assume "y" is always a consonant.

```{r}
length(grep("[^aeiou]{4}", a))
```

6. Return all words with a "q" that isn't followed by a "ui".

```{r}
grep("q[u?][^i]", a, value=TRUE)
```

7. Find all words that contain a "k" followed by another letter. Run the `table` command on the first character following the first "k" of each word.

```{r}
lst = grep("k[a-z]{1}", a, value=TRUE)

new = numeric(length(lst))
for (i in 1:length(lst)){
  
  letters = strsplit(lst[i], split="")
  idx = grep("k", letters)
  new[i] = letters[[1]][idx+1]
}

table(new)
```

8. Remove all vowels. How many character strings are found exactly once?

```{r}
a_new = numeric(length(a))
for (i in 1:length(a)){
  a_new[i] = gsub("[aeiou]", "", a[i])
}
length(unique(a_new))
```

### Question 3 ###

**3 points**

The first argument to most functions that fit linear models are formulas.  The following example defines the response variable `death` and allows the model to incorporate all other variables as terms. `.` is used to mean all columns not otherwise in the formula.

```{r}
url <- "https://github.com/couthcommander/Bios6301/blob/main/datasets/haart.csv"
tryCatch(haart_df <- read.csv(url)[,c('death','weight','hemoglobin','cd4baseline')], error=function(e) e)

haart_df = read.csv("~/Documents/Fall 2021/Statistical Computing/datasets/haart.csv")
coef(summary(glm(death ~ ., data=haart_df, family=binomial(logit))))
```

Now imagine running the above several times, but with a different response and data set each time.  Here's a function:

```{r}
myfun <- function(dat, response) {
  form <- as.formula(response ~ .)
  coef(summary(glm(form, data=dat, family=binomial(logit))))
}
```

Unfortunately, it doesn't work. `tryCatch` is "catching" the error so that this file can be knit to PDF.

```{r}
tryCatch(myfun(haart_df, death), error = function(e) e)
```

What do you think is going on?  Consider using `debug` to trace the problem.

(Note: I had to read in the data set from my local machine because the URL method was not working.) 

When I run `debug()` on `myfun`, I see that, although we specify that `response = death`, when `form` is created, it is created as `response ~ .` rather than `death ~ .` which makes me think there is an issue in how the formula is created. When I try to probe `response` it tells me `object 'death' not found`. Perhaps when we specify the response variable, we need to specify that it's a column of the data set, not some global object.

**5 bonus points**

Create a working function.


```

