---
title: 'Bios 6301: Assignment 8'
author: 'Haley Yaremych'
output: pdf_document
---

*Due Tuesday, 16 November, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

30 points total.

Submit a single knitr file (named `homework8.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework8.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**15 points**

Install the `readxl` package and run the following

```{r}
library(readxl)
fn <- 'icd10.xlsx'
if(file.access(fn, mode = 4) == -1) {
    url <- "https://www.cdc.gov/nhsn/xls/icd10-pcs-pcm-nhsn-opc.xlsx"
    download.file(url, destfile = fn, mode = 'wb')
}
dat <- readxl::read_excel(fn, sheet = 2)
```

1. Show the class of `dat`. (1 point)
```{r}
class(dat)
```

2. Show the methods available for objects of the given class (if there are multiple classes, show methods for all classes). (3 points)
```{r}
library(readxl)
library(dplyr)
tryCatch(methods(,tbl_df), error = function(e) e)
tryCatch(methods(,tbl), error = function(e) e)
tryCatch(methods(,data.frame), error = function(e) e)
```

3. If you call `print(dat)`, what print method is being dispatched? (1 point)

The print method being dispatched is a print function specifically for objects of the `tbl_df` class. 


4. Set the class of `dat` to be a data.frame. (1 point)
```{r}
dat = as.data.frame(dat)
```

5. If you call `print(dat)` again, what print method is being dispatched? (1 point)

Now, since `dat` is a type data.frame, the print method being dispatched is the print method that is specifically for objects of type `data.frame`, which could also be called with `print.data.frame`. 


Define a new generic function `nUnique` with the code below.

```{r}
nUnique <- function(x) {
    UseMethod('nUnique')
}
```

6. Write a default method for `nUnique` to count the number of unique values in an element. (2 points)
```{r}
nUnique.default = function(x){
  return(length(unique(x)))
}
methods("nUnique")
```

7. Check your function (2 points)

```{r}
nUnique(letters) # should return 26
nUnique(sample(10, 100, replace = TRUE)) # should return 10 (probably)
```

8. Write a data.frame method for `nUnique` to operate on data.frame objects.
This version should return counts for each column in a data.frame. (2 points)
```{r}
nUnique.data.frame = function(df){
  apply(df, MARGIN=2, FUN=nUnique)
}
```

9. Check your function (2 points)

```{r}
nUnique(dat)
```

### Question 2 ###

**15 points**

Programming with classes.  The following function will generate random patient information.

```{r}
makePatient <- function() {
  vowel <- grep("[aeiou]", letters)
  cons <- grep("[^aeiou]", letters)
  name <- paste(sample(LETTERS[cons], 1), sample(letters[vowel], 1), sample(letters[cons], 1), sep='')
  gender <- factor(sample(0:1, 1), levels=0:1, labels=c('female','male'))
  dob <- as.Date(sample(7500, 1), origin="1970-01-01")
  n <- sample(6, 1)
  doa <- as.Date(sample(1500, n), origin="2010-01-01")
  pulse <- round(rnorm(n, 80, 10))
  temp <- round(rnorm(n, 98.4, 0.3), 2)
  fluid <- round(runif(n), 2)
  list(name, gender, dob, doa, pulse, temp, fluid)
}
```

1. Create an S3 class `medicalRecord` for objects that are a list with the named elements `name`, `gender`, `date_of_birth`, `date_of_admission`, `pulse`, `temperature`, `fluid_intake`. Note that an individual patient may have multiple measurements for some measurements.  Set the RNG seed to `8` and create a medical record by taking the output of `makePatient`.  Print the medical record, and print the class of the medical record. (5 points)
```{r}
set.seed(8)

medicalRecord = function(x){
  class(x) = 'medicalRecord'
  attr(x, 'name') = x[[1]] 
  attr(x, 'gender') = as.character(x[[2]])
  attr(x, 'date_of_birth') = x[[3]]
  attr(x, 'date_of_admission') = x[[4]] 
  attr(x, 'pulse') = x[[5]]
  attr(x, 'temp') = x[[6]]
  attr(x, 'fluid_intake') = x[[7]]
  return(x)
}

rec = medicalRecord(makePatient())
print(rec)
```

2. Write a `medicalRecord` method for the generic function `mean`, which returns averages for pulse, temperature and fluids. Also write a `medicalRecord` method for `print`, which employs some nice formatting, perhaps arranging measurements by date, and `plot`, that generates a composite plot of measurements over time.  Call each function for the medical record created in part 1. (5 points)
```{r}
library(ggplot2)

mean.medicalRecord = function(x){
  ret = rep(0, 3)
    ret[1] = mean(attr(x, 'pulse'))
    ret[2] = mean(attr(x, 'temp')) 
    ret[3] = mean(attr(x, 'fluid_intake'))
  return(ret)
}

print.medicalRecord = function(x){
  atts = attributes(x)[-1]
  
  DOB = as.character(attr(x, 'date_of_birth'), format="%Y-%m-%d")
  
  DOA = attr(x, 'date_of_admission')
  for (i in 1:length(DOA)){
    DOA[i] = as.character(DOA[i], format="%Y-%m-%d")
  }
  DOA = paste(sort(DOA))
  
  df = data.frame(Attribute = c("Name:", "Gender:", "DOB:",
                                "Pulse:", "Temp:", "Fluid Intake:"),
                     Value=rep(0,length(atts)-1))
  
  df$Value[1] = attr(x, 'name')
  df$Value[2] = attr(x, 'gender')
  df$Value[3] = DOB
  df$Value[4] = list(attr(x, 'pulse'))
  df$Value[5] = list(attr(x, 'temp'))
  df$Value[6] = list(attr(x, 'fluid_intake'))
  
  print(df)
  cat("Admission Date(s): ")
  cat(DOA)
}

plot.medicalRecord = function(x){
  df = data.frame(date = attr(x, 'date_of_admission'), 
                  temp = attr(x, 'temp'), 
                  pulse = attr(x, 'pulse'))
  colors <- c("Temp" = "red", "Pulse" = "purple")
  ggplot(data=df, aes(x=date)) + 
     geom_line(aes(y=temp, color="Temp")) +
     geom_line(aes(y=pulse, color="Pulse")) +
     labs(color="Legend") + 
     scale_color_manual(values=colors) +
    theme(axis.title.y = element_blank())
}

mean(rec)
print(rec)
plot(rec)
```

3. Create a further class for a cohort (group) of patients, and write methods for `mean` and `print` which, when applied to a cohort, apply mean or print to each patient contained in the cohort. Hint: think of this as a "container" for patients.  Reset the RNG seed to 8 and create a cohort of ten patients, then show the output for `mean` and `print`. (5 points)
```{r}
set.seed(8)

# accepts a list of medicalRecords
cohort = function(x){
  new = vector(mode = "list", length = length(x))
  
  for (i in 1:length(x)){
    new[[i]] = x[[i]]
  }
  class(new) = 'cohort'
  return(new)
}

mean.cohort = function(x){
  t(sapply(x, FUN=mean))
}

print.cohort = function(x){
  for (i in 1:length(x)){
    print(x[[i]])
    cat("\n--------\n")
  }
}

pats <- vector(mode = "list", length = 10)
for (i in 1:10){
  pats[[i]] = medicalRecord(makePatient())
}

cohort = cohort(pats)

mean(cohort)
print(cohort)

```




